---
title: "Importance of Data Visulization With Uber Dataset"
author: "Iman Jowkar"
date: "6/9/2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    math: katex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<span style="color:blue">In this project, I want to show you The Importance of data visualization.<br/>
Imagine you are a data scientist in an online taxi company, and your manager has a problem with three questions:<br/>
</span>.
<span style="color:red"> 1 - what times in a day the peak demand occurs?<br/>
2 - what days in a week the peak demand occurs?<br/>
3 - what days in a month the peak demand occur?<br/>
</span>.
<span style="color:blue">Your manager wants you to ask these questions by a dataset that gives you.<br/>
Your job is to analyze this data for answering these questions.<br/>
Ok, let's read the data set into R-studio:

</span>.

```{r echo = FALSE}

# clear work directory --------------------------

rm(list = ls())

# setwork directory --------------------------
setwd('F:/Fradars/ggplot2/2 - Uber Data Analysis with ggplot2/')


```

```{r import dataset}




# Read data------------------------

Jul_data <- read.csv("D:/uber-raw-data-jul14.csv", header = T)
Aug_data <- read.csv("D:/uber-raw-data-aug14.csv", header = T)
Sep_data <- read.csv("D:/uber-raw-data-sep14.csv", header = T)



# View(Jul_data)
# View(Aug_data)
# View(Sep_data)

str(Jul_data)

```

<span style="color:blue"> We have 4th columns(feature) in these datasets. 
we only work with the specific column called 'Date.Time'.
Lets adding some column to answer these questions
</span>.



```{r adding some column}
# First of all we merge our datasets


data = rbind(Jul_data, Aug_data, Sep_data)

# change the format of 'Date.Time' column to time-format

data$Date.Time <- as.POSIXct(data$Date.Time, format = "%m/%d/%Y %H:%M:%S")
class(data$Date.Time)
# str(data)
# View(data)


# Lets add a net column called 'Time'
data$Time <- format(as.POSIXct(data$Date.Time, format = "%m/%d/%Y %H:%M:%S"),
                         format="%H:%M:%S")

# View(data)


library(lubridate) # Our dataset involves various time-frames.
# In order to understand our data in separate time categories,
# we will make use of the lubridate package.
data$day = factor(day(data$Date.Time))
data$month = factor(month(data$Date.Time, label = T))
# data$year = factor(year(April_data$Date.Time))
data$dayofweek = factor(wday(data$Date.Time, label = T, abbr = F))
data$hour = factor(hour(hms(data$Time)))
# View(data)
head(data)
```

```{r }
# Hourly Demand
hour_data = as.data.frame(table(data$hour))
names(hour_data) = c('Hour', 'Demand')
head(hour_data)
# View(hour_data)


library(ggplot2)
library(ggthemes)
library(scales)
ggplot(hour_data, aes(x = Hour, y = Demand, group = 1)) + 
  geom_bar(stat = 'identity', fill = 'blue',col = 'white') +
  geom_point() +
  scale_y_continuous(labels = comma) + 
  annotate("text", label = "Pick Demand", x = 17, y = 198000, color = "red", size = 6) +
  ylab('Number of Demand') +
  theme_clean() + 
  ggtitle('Hourly Demand')

# It seems that about 17 and 18 we have pick demand.



```
```{r }
hour_month_demand = table(data$hour, data$month)
hour_month_demand
class(hour_month_demand)
hour_month_demand = as.data.frame(hour_month_demand)
# View(hour_month_demand)
names(hour_month_demand) = c('Hour','Month','Demand')

ggplot(hour_month_demand, aes(x = Hour, y = Demand, fill = Month)) + 
  geom_bar(stat = 'identity') +
  scale_y_continuous(labels = comma) + 
  annotate("text", label = "Pick Demand", x = 17, y = 198000, color = "red", size = 6) +
  theme_base() + 
  ggtitle('Hourly and monthly Demand')

library(dplyr)

July = data %>% 
  filter(month == 'Jul')

Augest = data %>%
  filter(month == 'Aug')

September = data %>% 
  filter(month == 'Sep')

hourly_demand_July = as.data.frame(table(July$hour))
names(hourly_demand_July) = c('Hour','Demand')
head(hourly_demand_July)

p1 = ggplot(data = hourly_demand_July, aes(x = Hour, y = Demand)) + 
  geom_bar(stat = 'identity', fill = 'blue', col = 'red') + 
  ggtitle('July Hourly Demand') + 
  theme_bw() + 
  scale_y_continuous(breaks = seq(0,70000,by = 10000)) + 
  expand_limits(y = c(0, 70000)) + 
  geom_hline(yintercept=35000, linetype="dashed", color = "red")

p1

hourly_demand_Augest = as.data.frame(table(Augest$hour))
names(hourly_demand_Augest) = c('Hour','Demand')
head(hourly_demand_Augest)

p2 = ggplot(data = hourly_demand_Augest, aes(x = Hour, y = Demand)) + 
  geom_bar(stat = 'identity', fill = 'blue', col = 'red') + 
  ggtitle('Augest Hourly Demand') + 
  theme_bw() + 
  scale_y_continuous(breaks = seq(0,70000,by = 10000)) +
  expand_limits(y = c(0, 70000)) + 
  geom_hline(yintercept=35000, linetype="dashed", color = "red")

p2


hourly_demand_September = as.data.frame(table(September$hour))
names(hourly_demand_September) = c('Hour','Demand')
head(hourly_demand_September)

p3 = ggplot(data = hourly_demand_September, aes(x = Hour, y = Demand)) + 
  geom_bar(stat = 'identity', fill = 'blue', col = 'red') + 
  ggtitle('September Hourly Demand') + 
  theme_bw() + 
  scale_y_continuous(breaks = seq(0,70000,by = 10000)) + 
  expand_limits(y = c(0, 70000)) + 
  geom_hline(yintercept=35000, linetype="dashed", color = "red")

p3


plot1 = gridExtra::grid.arrange(p1, p2,p3, nrow = 3)
# By seeing The graphs, we can answer question one:
# As for the solution, I think It seems that the peak demand every three months occurs about 17 and 18.
```


```{r weekly demmand}
# weekly demand

Weekly_demand = as.data.frame(table(data$dayofweek))
head(Weekly_demand)
names(Weekly_demand) = c('Day','Demand')
Weekly_demand

red.bold.italic.text <- element_text(face = "bold.italic", color = "red", size = 16)
black.bold.italic.16.text <- element_text(face = "bold.italic", color = "black", size = 10)
ggplot(data = Weekly_demand, aes(x = Day, y = Demand)) + 
  geom_bar(stat = 'identity', fill = 'blue', col = 'red') + 
  ggtitle('Weekly Demand') + 
  theme_bw() + 
  theme(title = red.bold.italic.text, axis.title = red.bold.italic.text) + 
  theme(axis.text = black.bold.italic.16.text)


Week_Month_Demand = as.data.frame(table(data$dayofweek, data$month))
head(Week_Month_Demand)
names(Week_Month_Demand) = c('Day','Month','Demand')

ggplot(data = Week_Month_Demand, aes(x = Day, y = Demand, fill = Month)) + 
  geom_bar(stat = 'identity') + 
  ggtitle('Weekly Demand') + 
  theme_bw()



Week_Demand_Augest = as.data.frame(table(Augest$dayofweek))
head(Week_Demand_Augest)
names(Week_Demand_Augest) = c('Day','Demand')

p1 = ggplot(data = Week_Demand_Augest, aes(x = Day, y = Demand)) + 
  geom_bar(stat = 'identity', fill = 'blue', col = 'red') + 
  ggtitle('Augest Weekly Demand') + 
  theme_bw()

p1



Week_Demand_July = as.data.frame(table(July$dayofweek))
head(Week_Demand_July)
names(Week_Demand_July) = c('Day','Demand')

p1 = ggplot(data = Week_Demand_July, aes(x = Day, y = Demand)) + 
  geom_bar(stat = 'identity', fill = 'blue', col = 'red') + 
  ggtitle('July Weekly Demand') + 
  theme_bw()


Week_Demand_Augest = as.data.frame(table(Augest$dayofweek))
head(Week_Demand_Augest)
names(Week_Demand_Augest) = c('Day','Demand')

p2 = ggplot(data = Week_Demand_Augest, aes(x = Day, y = Demand)) + 
  geom_bar(stat = 'identity', fill = 'blue', col = 'red') + 
  ggtitle('Augest Weekly Demand') + 
  theme_bw()



Week_Demand_september = as.data.frame(table(September$dayofweek))
head(Week_Demand_september)
names(Week_Demand_september) = c('Day','Demand')

p3 = ggplot(data = Week_Demand_september, aes(x = Day, y = Demand)) + 
  geom_bar(stat = 'identity', fill = 'blue', col = 'red') + 
  ggtitle('September Weekly Demand') + 
  theme_bw()


plot2 = gridExtra::grid.arrange(p1, p2,p3, nrow = 3)

```


```{r monthly demand}
Montly_Demand = as.data.frame(table(data$month))
head(Montly_Demand)
names(Montly_Demand) = c('Month','Demand')

p = ggplot(data = Montly_Demand, aes(x = Month, y = Demand)) + 
  geom_bar(stat = 'identity', col = 'red', fill = 'blue') + 
  theme_bw()

p
# It seems That in September The company has a big demand rather than the previous month.
```


As for The results:<br>
question one: Hourly demand on each day is about 17 & 18.<br>
question Two: This is not an obvious trend in each month.<br>
question one: on september, we have a higher demand rather than the other month.


